What can GitHub and Googlecode Issues tell us about Exceptions in Android and Java?

### 2014-09-01 android-stacks.tex
all over . K9mail

Different from GitHub, Googlecode does not provide an API to access the information related
 to hosted projects \footnote{Googlecode previoulsly provided a Web service but was deactivated in June 2013.}.
However many well used open source Android application such as K9mail, is hosted in Googlecode which stores
the issue history of this and other apps are store there ~\footnote{K9mail moved to Github but advice their 
users to create issues on Googlecode as a way of not loosing the issue history.}.

To overcome this limitation and mine the information available on such Googlecode 
projects we implemented a Web crawler which navigated (incorporated in Exception Miner tool described next) 
 on each project and issue page, extracting the content of each issue to store on relational database. \footnote{Google code previously provided a Web service to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}

To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navigated 
through the web interface of Googlecode projects extracting all issues and issue comments and storing in a relational database for later analysis.

@@

@@

\emph{\textbf{Direct instances of RuntimeExceptions being thrown.}}
From Table ~\ref{tab:topten} we can observe that direct instances
java.lang.RuntimeException were thrown in 18,88\% of Android repositories, and
most of such exceptions  were  thrown by the Android platform (167 out of 232).

Throwing general exceptions, such as direct instances of RuntimeException, violates best
practice III. It is considered a bad practice because exception type does not carry enough
 information to identify the cause of the exceptional behavior, and as a consequence developers need to relying on
the exception message which can be neither complete nor precise~\cite{gosling2000java}.

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
Programming mistakes were reported as the root cause of  52\% of the analyzed stack traces. 
In 30,08\% NullPointerExceptions was reported as the root cause.
}
\end{minipage}}

@@

Besides distilling the information available on stack traces (i.e., exception signalers, root causes, exception wrappings),
we mined the exception types (i.e., RuntimeException, checked Exception or Error) related to each exception reported on stack traces 
using source code and byte code analyses supported by Exception Miner described before.

 Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on the analyzed issues. As we can see 
from Table ~\ref{tab:typeroottab},  most of the reported exceptions were signaled by Android and libraries reused by Apps were runtime exceptions. 
On the other hand, we could observe that most of the reported exceptions signaled by the set of 
libraries reused by Android platform (i.e., org.apache.harmony,
org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax -- referred in
this work as libcore) were checked exceptions. It is pointed as a good practice
for libraries (see Section~\ref{sec:best}) because by using checked exceptions
libraries can define a precise exception interface to its clients. Such
libraries are heavily used in several projects, and such precise exception
interface may be related to the library's maturity.

@@

We could observe that most of the exceptions reported on issues were runtime, 
mainly  signaled on the Android Platform and in the Application code. we could also find 
runtime exceptions signaled from libraries.

In order to evaluate if  best practice IV was being obeyed, we manually inspected the   
signature of each method signaling a runtime exception, that was defined by third 
party code (lib or Android) and reused by an application.
In this task we filtered out all the exceptions implicitly 
signaled by  JVM (due to programming mistakes) since these exceptions should not be documented on the 
method signature.

We could observe that only 1 method out of 118 inspected signalers
 (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other  study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.

@@

We could also observe checked exceptions (which according to Java specification represent
recoverable conditions) as the root causes of reported exception stack traces.
Two possible ways of a checked exception can remain uncaught are: the checked 
exception is declared by every method on the call chain or if it is wrapped in a 
runtime exception and re-thrown. However the inspection of stack traces 
and the signatured of methods revealed the a third option: an undocumented
checked exception thrown by a native method.

The code snippet in Figure ~\ref{fig:snippets} (d) illustrate this scenario.
 The java/lang/Class declares a native method named getDeclaredMethods. 
The Java-side declaration of of this method does not have any throws clause, 
leading programmers and the compiler to think that no checked exceptions can be thrown.

 However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
violating the declaration. The Java compiler, could not detect this violation because it does 
not perform static exception checking on native methods. This type of bug is hard to diagnose
because the developer usualy do not have access to the native implementations.

Consequently, since it is not expected by the programmer, when such method throws 
this exception, such undocumented exception may remain
uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
The exception stack traces reporting this scenario is actually a real bug of Android 
Gingerbread version (which still accounts for 13.6\% of devices running Android).

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
 Only 0,8\% of the methods documented the explicitly signaled runtime exceptions.
  The study also detected an undocumented checked exception signaled by
  a native method defined in Android.
}
\end{minipage}}

\subsection{What can the Exception Wrappings tell us?}
%\subsection{Exploring the Information Embedded on Exception Wrappings}
Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions were indeed wrapped in runtime exceptions or even errors.

%, while most of them were not wrapped along the stack trace.

@@

\begin{table*}
  \centering
  \begin{tabular}{llcccccc}
    \hline
    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{Projects}  &  \bfseries{Occurrences} & \textsf{Android} & \textsf{Java/Libcore} & \textsf{Lib} & \textsf{App}  \\
    \hline
      %Runtime &  Checked   & 70  / 18 & 114 /  34 &  66 /  9  &  0 /  & 19 / 19 &  29 / 6 \\

      Runtime &  Checked   & 88 & 148 &  75  &  0   & 38 &  35 \\

      %Runtime   &  Error   & 38  \ 8  & 57    &  50 \ 8  &  0   & 6 \ 2  &  1   /  0\\
      Runtime   &  Error   & 46  &  67    &  58  &  0   & 8  &  1   \\

%      Checked &  Runtime   & 12 \ 5  & 22 \ 9 & 4 &  0  & 9 \7 &  9 \ 2 \\
      Checked &  Runtime   & 17  & 31 & 4 &  0  & 16 &  11 \\

%      Checked & Error      & 7 \ 1 &  8 & 5  &  0  & 0 \ 1 &  3 \0 \\

      Checked & Error      & 8 &  9  & 5  &  0  &  1 &  3  \\

 %     Error & Checked      & 4 \ 10  &  7 & 0  \ 6  &  7  & 0 \6 &  0   \ 8    \\
     Error & Checked      & 14 &  27 &  6  &  7  &  6 &   8    \\

%      Error & Runtime     & 6 \ 2  &  13   & 1 &  1  & 0 \ 1 &  11 \ 3    \\

      Error & Runtime     & 8  &  17   & 1 &  1  & 1 &  14    \\

@@

TODO: ATUALIZAR DISCUSSAO COM BASE NOS NOVOS DADOS.

This analysis revealed unexpected wrappings such as: Error wrapping a Checked 
exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
Java is the only language that provides a hybrid exception handling mechanism
which offers  different types to represent different exception behaviors (i.e., error, 
runtime and checked) (see Section~\ref{sec:best}). According to Java specification Errors should not be
handled inside the system since they usually represent unrecoverable conditions
detected by the JVM such as OutOfMemoryError. Checked exceptions should
represent recoverable conditions and  Runtime exceptions represent are usually
related to programming mistakes from which the developer is not expected to recover
from. Next sections discusses about such types of wrappings.

\emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
detected on 70 distinct projects. Inspecting the 
elements responsible for performing such wrapping, we could observe that
most of such wrappings were performed on methods defined on Android
Platform (approximately 58%).

%\begin{table}
%\centering
%\begin{tabular}{ll}
%    \hline
%   \bfseries{signaler from} & \bfseries{Ocurrences} \\
%    \hline
%ANDROID	 & 66 \\
%APP	& 29 	(HOW MANY?? AND WHICH??) \\
%LIB  & 19 (HOW MANY AND WHICH?)\\
%\hline
% \end{tabular}
%\caption{Elements responsible for the wrapping.}
%\label{tab:wrapping01}
%\end{table}

Taking a closer look on the methods responsible for such wrappings,
 we could observe that approximately 41% of them were performed by a single class,
 named ActivityThreat. This class is responsible for managing the life cycle of an Android Activities.

\begin{table}
\centering
\begin{tabular}{lll}
    \hline
 \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
    \hline
android.app.ActivityThread & java.lang.RuntimeException & 47 \\
android.view.LayoutInflater & android.view.InflateException &  7 \\
android.os.AsyncTask & java.lang.RuntimeException &  5 \\
Other platform classes & java.lang.RuntimeException &   7 \\
\hline
  \end{tabular}
\caption{Elements tresponsible for the wrapping.}
\label{tab:wrapping01}
\end{table}

Figure ~\ref{fig:snippets} (a) presents a code snippet of ActivityThreat which is responsible for (among other things)
calling the Activities onCreate method. Hence, as we can see from lines 4-8,
 in case any exception is thrown by an application Activity during its creation, 
it is wrapped in a general Runtime exception and re-thrown.

@@

The inversion of control makes it difficult to the application developer to associate 
a specific handler to the exceptions thrown during the onCreate method (or any other Activity life cycle method).
Since according to Java handling model the handler should me added on any method in the call chain,
and in the inversion of control development model the developer usually does not have fully access to 
the invocation call chain of a method in order to associate a handler, as a result the 
exception may scape from any handling and remain uncaught.

@@

\emph{\textbf{Runtime Exception wrapping Error.}} We could find  57 stack traces 
on 38 projects on which an Error was wrapped in a runtime exception.
Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
The code snippet below shows that any throwable signaled in the context of an asynchronous 
task is converted in to a runtime and re-thrown.

ASSNCTASKS NON ANNOTATED METHODS…….
SOME HAVE EXCEPTION INTERFACE 
EXCEPTIONS ON MEHTODS SIGNATURE OTHERS NOT
INCONSISTENCY…. DEPENDS ON DEVELOPERS 
WORK INSTEAD OF TOOLS.

@@

We could observe that several types of exceptions were encapsulated in 
instances of Runtime such as java.lang.OutOfMemoryError (30), java.lang.NoClassDefFoundError (6), 
java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}.

@@

\emph{\textbf{Checked wrapping Runtime Exception.}} Manually inspecting the stack 
traces we could observe that most of the wrappings were performed 
by RoboSpice (an android library that supports the creation of asynchronous long running tasks.
This library has 1.682 stars and 367 forks) and by the reflection Java library (used by Android, 
Application and library classes). Inspecting the code related to such wrappings we could observe that
the reflection library convert every exception that happens 
when invoking a reflected method into an InvoncationTargetException (this conversion is 
performed inside native code).

@@

\emph{\textbf{Checked Exception wrapping Error.}} Almost all of these wrappings
 were also caused by the reflection library 
used by Android applications. The method responsible for the wrappings
were native methods written in C. Table  ~\ref{tab:wrapcheckerr} illustrates
some oth these wrappings some of them are masking an OutOfMemoryError
into a checked exception.

@@

\emph{\textbf{Error wrapping Runtime.}} After manually inspecting the code associated to such wrappings we discovered that all of them were 
caused by static initializers. If any exception is thrown in the context of a static initializer (i.e., static block) 
it is converted into an ExceptionInitializerError on the point where the class is first used. Table X illustrates
the exception types involved in such wrappings.

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError)
reveal fragilities on Java Exception Model.
}
\end{minipage}}

@@

@@

When family of systems make use of a common core, or framework such systems can
benefit from a cross project crash analysis. The bug issues reported for a group of systems
can identify vulnerabilities which were more difficult to identify in a single project analysis.

The test case explosion problem... A test suite is always incomplete,
one application could benefit from the usage scenarios on which such framework is applied.
Framework testing -  what are the main challenges on framework testing?
A cross-project crash analysis based on repository mining techniques can 
identify vulnerabilities in pieces of code that are shared.

Using exception miner tool we could identify the methods responsible for most of the crashes - 
such methods can be a target for manual inspections improvements, or better documentation
(because sometimes an exception can be thrown because the user is not invoking the method
in the proper way ....

@@

It allows the developers to benefit from the faults happening on different systems.

@@

\noindent\emph{External Validity} The results presented here were based on mining
issues on Github through the GHTorrent dataset. While comprehensive and
extensive, GHTorrent is not an exact replica of Github, so several issues might
be left out. Due to the way data collection works with GHTorrent, for projects
that are relatively inactive, GHTorrent might have not collected a significant
proportion of their issues. We did not investigate the extend of this threat on
our sample.
<OLHAR ARTIGO DO FSE QUE TRABALHA COM OPENSOURCE.
FALAR QUE DESCOBERTAS FORAM DE LIBS JAVA. APESAR DO ESTUDO NAO SER FEITO EM PROJS JAVA.....
PODEMOS CONCLUIR ALGUMAS COISAS QUE SE APLICAM A TODOS JAVA>

@@

\textit{Extracting Stack Traces from natural language artifacts.} 
%Currently, many
%software vendors embed automatic crash reporting tools in their software
%systems. Moreover, third party crash collection services exist, most of them 
%targeted for applications run on mobile phones~\cite{BugSe14,BugSn14,Googl14,Acra14}.
Apart from bug reports, stack traces can be embedded in other forms of
communication between developers, such as discussion logs and emails.
Being intermixed with text makes the accurate extraction of stacktraces 
an involved process.
Infozilla~\cite{bettenburg2008extracting}
is based on a set of regular expressions that extract a set of frames
related to a stack trace. The main limitation of this solution is that it is not
able to extract stack traces embedded on verbose log files (i.e., on which we
can find log text mixed with exception frames). Bacchelli
et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
from development emails and relate it to code artifacts (i.e. classes) mentioned
on the stack trace. In addition to those tools, ExceptionMiner is able to 
both extract stack traces from natural language artifacts and to 
classify them in a set of predefined categories.
(MENCIONAR QUE X% DOAS EXCECOES ANALISADAS ESTAVAM EMBEBIDAS EM ALGUM TIPO DE ARQUIVO DE LOG - 
LOGCAT DO ANDROID....)

@@

Additionally, Cabral and Marques~\cite{cabral2007exception} analyzed the
exception handling code of 32 open-source systems, both for Java and .NET. They
observed that the action handlers were very simple (e.g., logging and present a
message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
set of bad practices on exception handling that hinder software maintainability
and robustness, based on their own experience with Java enterprise applications.
Our work differs from those two works as we tried to identify bad practices from
the analysis of stack traces extracted from issues in combination with and bytecode and source
code analysis.

RESSALTAR LIMITACAO DE QUANTIDADE DE FALSOS POSITIVOS... E QUE OS STACKS SAO REAIS...
REPRESENTAM FLUXOS REAIS.... POST MOSTERN ANALISYS OF REAL EXECUTIONS....

>>>>

Differently from GitHub, Googlecode does not provide an API to access the information related
 to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.

However, Googlecode still contain largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
it advices their users to report bugs on Googlecode issue tracker.})  and the history of such apps are also valuable. 
To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navigated 
through the web interface of Googlecode projects extracting all issues and issue comments and storing in a relational database for later analysis.

@@

%For instance: K9mail  was hosted in Googlecode 
%before moving to GitHub and in order not to loose the project history it advices their 
%users to create issues on Googlecode as a way to preserve project history.

%To overcome this limitation and mine the information available on such Googlecode 
%projects we implemented a Web crawler which navigated (incorporated in Exception Miner tool described next) 
% on each project and issue page, extracting the content of each issue to store on relational database.

@@

%\emph{\textbf{Direct instances of RuntimeExceptions being thrown.}}
%From Table ~\ref{tab:topten} we can observe that direct instances
%java.lang.RuntimeException were reported as the root cause of stacks in 19,09\% 
%of the projects, and most of such exceptions  were  thrown by the Android platform (167 out of 232).
%Throwing general exceptions, such as direct instances of RuntimeException, violates best
%practice III. It is considered a bad practice because exception type does not carry enough
%information to identify the cause of the exceptional behavior, and as a consequence developers need to relying on
%the exception message which can be neither complete nor precise~\cite{gosling2000java}.

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
A multitude of programming mistakes were reported as the main causes of stack traces.
51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.
}
\end{minipage}}

@@

Besides distilling the information available on stack traces (i.e., exception signalers, root causes, exception wrappings),
we mined the exception types (i.e., RuntimeException, checked Exception or Error) related to each exception reported on stack traces -
to do so we used the source code and byte code analyses implemented on Exception Miner.

 Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on the analyzed issues. As we can see 
from Table ~\ref{tab:typeroottab},  most of the reported exceptions signaled (explicilty or implicitly by the JVM) on methods defined
 on the Android Plaform on on the application itself.

On the other hand, most of the reported exceptions signaled by the set of 
libraries reused by Android platform were checked exceptions. Such libraries are referred in
this work as libcore and is composed by org.apache.harmony, org.w3c.dom, sun.misc, 
org.apache.http, org.json, org.xml, javax. It is pointed as a good practice
for libraries (see Section~\ref{sec:best}) because by using checked exceptions
a library can define a precise exception interface ~/cite{miller1997issues} to its clients. Such
libraries are heavily used in several projects, and such precise exception
interface may be related to the library's maturity.

@@

\emph{\textbf{Undocumented Runtime Exceptions}}.
However, we can also observe a considetable amount of 
runtime exception signaled on library reused by apps.
In order to evaluate if  best practice IV was being obeyed, we inspected the   
signature of each method signaling a runtime exception. 
We inspected methods of the Android Plaftorm and 
the libraries, because both are  third  party code reused 
by the Android apps. In this task we filtered out all the exceptions implicitly 
signaled by  JVM (due to programming mistakes) 
since these exceptions should not be documented on the 
method signature.

As a result of this inspection, we could observe that only 1 method out of 118 inspected signalers
 (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.

@@

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
 Only 0,8\% of the methods documented the explicitly signaled runtime exceptions. 
}
\end{minipage}}

\subsection{What can the Exception Wrappings tell us?}
%\subsection{Exploring the Information Embedded on Exception Wrappings}

Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions were indeed wrapped in runtime exceptions or even errors.

@@

 %Runtime &  Checked   & 70  / 18 & 114 /  34 &  66 /  9  &  0 /  & 19 / 19 &  29 / 6 \\
      %Runtime   &  Error   & 38  \ 8  & 57    &  50 \ 8  &  0   & 6 \ 2  &  1   /  0\\
%      Checked &  Runtime   & 12 \ 5  & 22 \ 9 & 4 &  0  & 9 \7 &  9 \ 2 \\
%      Checked & Error      & 7 \ 1 &  8 & 5  &  0  & 0 \ 1 &  3 \0 \\
 %     Error & Checked      & 4 \ 10  &  7 & 0  \ 6  &  7  & 0 \6 &  0   \ 8    \\
%      Error & Runtime     & 6 \ 2  &  13   & 1 &  1  & 0 \ 1 &  11 \ 3    \\

\begin{table*}
  \centering
  \begin{tabular}{llcccccc}
    \hline
    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{Projects}  &  \bfseries{Occurrences} & \textsf{Android} & \textsf{Java/Libcore} & \textsf{Lib} & \textsf{App}  \\
    \hline

      Runtime &  Checked   & 88 & 148 &  75  &  0   & 38 &  35 \\
      Runtime   &  Error   & 46  &  67    &  58  &  0   & 8  &  1   \\      
      Checked &  Runtime   & 17  & 31 & 4 &  0  & 16 &  11 \\
      Checked & Error      & 8 &  9  & 5  &  0  &  1 &  3  \\
      Error & Checked      & 14 &  27 &  6  &  7  &  6 &   8    \\
      Error & Runtime     & 8  &  17   & 1 &  1  & 1 &  14    \\

@@

This analysis revealed unexpected wrappings such as: Error wrapping a Checked 
exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
Java is the only language that provides a hybrid exception model 
which offers  different types to represent different exception behaviors (i.e., error, 
runtime and checked) (see Section~\ref{sec:best}). For instance,
according to Java specification, Errors should not be handled inside the system 
since they usually represent unrecoverable conditions
detected by the JVM (e.g.,  OutOfMemoryError). On the other hand, checked exceptions should
represent recoverable conditions. Next we discuss and exemplify the cross-type wrappings
 found in this study.

\emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
detected on 88 projects. Inspecting the 
elements responsible for performing such wrapping, we could observe that
approximately  50\% of such wrappings were performed on methods defined on Android
Platform.

Taking a closer look on the methods responsible for such wrappings,
 we could observe that approximately 41\% of them were performed by a single class,
 named ActivityThreat.

This class is responsible for managing the life cycle of  Android Activities - 
invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
(window) of an Android app.
Figure ~\ref{fig:snippets} (a) presents a code snippet of one the methods of 
ActivityThrea,  the one responsible for invoking the Activities onCreate method. 
As we can see from lines 4-8,
 in case any exception is thrown in the context of such try-catch block, 
it is wrapped in a general Runtime exception and re-thrown.

\begin{table}
\centering
\begin{tabular}{lll}
    \hline
 \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
    \hline
android.app.ActivityThread & java.lang.RuntimeException & 61\\
android.view.LayoutInflater & android.view.InflateException &  7 \\
android.os.AsyncTask & java.lang.RuntimeException &  7\\
\hline
  \end{tabular}
\caption{Top 3 Runtime-Cheked wrappers.}
\label{tab:wrapping01}
\end{table}

@@

%SELECT SIGNALER_method, COUNT(*) AS N
%  FROM stacktracesummary WHERE ROOT_EXTENDS='CHECKED' AND EXCEPTION_EXTENDS='RUNTIME' and
%  ( SIGNALER_method like '%onDestroy%' or
%   SIGNALER_method like '%onCreate%' or
% SIGNALER_method like '%onStart%' or
% SIGNALER_method like '%onResume%' or
% SIGNALER_method like '%onPause%' or
% SIGNALER_method like '%onStop%' or
% SIGNALER_method like '%onDestroy%' )
%   GROUP BY SIGNALER_method
%  ORDER BY N DESC;

%SELECT COUNT(*) AS N
 % FROM stacktracesummary WHERE ROOT_EXTENDS='CHECKED' AND EXCEPTION_EXTENDS='RUNTIME' and
 % ( stack like '%onDestroy%' or
 %  stack like '%onCreate%' or
% stack like '%onStart%' or
% stack like '%onResume%' or
% stack like '%onPause%' or
% stack like '%onStop%' or
% stack like '%onDestroy%' );

%GC = 6

However, since the Activity life clycle method (i.e., callActivityOnCreate(...)) that is invoked in this block,
 do not include checked exceptions on its signatures. Hence they could not throw a checked exception to be handled 
in the method illustrated above. However the code inspection performed on such wrappings revelaed an
``undocumented checked exception" signaled by this life cycle method.

\emph{\textbf{Undocumented Checked Exceptions.}} There are usually
2 possible possibilities for a checked exception to be thrown by a method:
it can either be specified on the method signature or it can be wrapped in a 
runtime exception and re-thrown. However by inspectioning associated to the reported wrappings 
revealed the a third option: a checked exception thrown by a native method, which was not declared
on the method signature of the methods on the call chain composing the stack.

The code snippet in Figure ~\ref{fig:snippets} (d) illustrate this scenario.
 The java/lang/Class declares a native method named getDeclaredMethods. 
The Java-side declaration of of this method does not have any throws clause, 
leading programmers and the compiler to think that no checked exceptions can be thrown.
 However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
violating the declaration. The Java compiler, could not detect this violation because it does 
not perform static exception checking on native methods. This type of bug is hard to diagnose
because the developer usualy do not have access to the native implementations. 
Consequently, since it is not expected by the programmer, when such method throws 
this exception, such undocumented exception may remain
uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
The exception stack traces reporting this scenario is actually a real bug of Android 
Gingerbread version (which still accounts for 13.6\% of devices running Android).

@@

\emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
on 46 projects on which an Error was wrapped in a runtime exception.
Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any throwable signaled in the context of an asynchronous 
task is converted in to a runtime and re-thrown. Some of the Erros wrapped on instances of runtime exceptions were 
java.lang.OutOfMemoryError (30), java.lang.NoClassDefFoundError (6), 
java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}

%ASSNCTASKS NON ANNOTATED METHODS…….
%SOME HAVE EXCEPTION INTERFACE 
%EXCEPTIONS ON MEHTODS SIGNATURE OTHERS NOT
%INCONSISTENCY…. DEPENDS ON DEVELOPERS 
%WORK INSTEAD OF TOOLS.

@@

.

@@

\emph{\textbf{Checked wrapping Runtime Exception.}} Manually inspecting the stack 
traces we could observe that most of the wrappings were performed 
by RoboSpice (an android library that supports the creation of asynchronous long running 
tasks, which has 1.682 stars and 367 forks in GitHub) and by the reflection Java library (used by Android, 
Application and library classes). Inspecting the code related to such wrappings we could observe that
the reflection library convert every exception that happens 
when invoking a reflected method into an InvoncationTargetException (this conversion is 
performed inside native code).

@@

\emph{\textbf{Checked Exception wrapping Error.}} Almost all of these wrappings
 were also caused by the reflection library 
used by Android applications. The methods responsible for the wrappings
were also native methods written in C. Table ~\ref{tab:exampleswrap} illustrates
some oth these wrappings some of them are masking an OutOfMemoryError
into a checked exception.

@@

\emph{\textbf{Error wrapping Runtime.}} After manually inspecting the code associated to such wrappings 
we discovered that all of them were caused by static initializers. If any exception is thrown in the context of a static initializer (i.e., static block) 
it is converted into an ExceptionInitializerError on the point where the class is first used. Table ~\ref{exampleswrap} illustrates
the exception types involved in such wrappings.

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError)
reveal fragilities on Java Exception Model.
The study also detected an undocumented checked exception signaled by
  a native method defined in Android.
}
\end{minipage}}

@@

\subsection{Inversion of Control and Exception Handling}
Reasoning about the design strategy adopted by the ActivityThread,
we can observe that the inversion of control (on which this class and 
most of the Android framework is based) can make it difficult for the application developer to 
associate a specific handler to the exceptions thrown during the execution of the onCreate method (or any other life cycle method).
According to Java handling model, the handler should be added on a method on the invocation chain. 
However,  in the inversion of control development model the developer does not have fully access to 
the invocation chain of a method. Therefore, the developer may be tempted to convert the checked
 exception into a RuntimeException and resignal.

@@

When family of systems make use of a common core, it is possible to benefit from a cross project crash analysis like the one performed in this study.
 The not only the exception stack traces but any other bug issues reported for a group of systems can identify vulnerabilities 
that are common to more then one element of the family and are more difficult to identify in a single project analysis.
Therefore tools such as the one developed in this study may allow the developers to 
benefit from the faults happening on different systems.

%It can be a way of dealing with the test case explosion problem related to framework test.
%Let the products test and report the crahes in a single repository and then analize them.
%... A test suite is always incomplete,
%one application could benefit from the usage scenarios on which such framework is applied.
%Framework testing -  what are the main challenges on framework testing?

%Using exception miner tool we could identify the methods responsible for most of the crashes - 
%such methods can be a target for manual inspections improvements, or better documentation
%(because sometimes an exception can be thrown because the user is not invoking the method
%in the proper way ....

@@

@@

\noindent\emph{External Validity} The results presented here were based on mining
issues on Github through the GHTorrent dataset. While comprehensive and
extensive, GHTorrent is not an exact replica of Github, so several issues might
be left out. Due to the way data collection works with GHTorrent, for projects
that are relatively inactive, GHTorrent might have not collected a significant
proportion of their issues. We did not investigate the extend of this threat on
our sample.
\footnote{to do: OLHAR ARTIGO DO FSE QUE TRABALHA COM OPENSOURCE.
FALAR QUE DESCOBERTAS FORAM DE LIBS JAVA. APESAR DO ESTUDO NAO SER FEITO EM PROJS JAVA.....
PODEMOS CONCLUIR ALGUMAS COISAS QUE SE APLICAM A TODOS JAVA}

@@

\textit{Extracting Stack Traces from natural language artifacts.} 
%Currently, many
%software vendors embed automatic crash reporting tools in their software
%systems. Moreover, third party crash collection services exist, most of them 
%targeted for applications run on mobile phones~\cite{BugSe14,BugSn14,Googl14,Acra14}.
Apart from bug reports, stack traces can be embedded in other forms of
communication between developers, such as discussion logs and emails.
Being intermixed with text makes the accurate extraction of stacktraces 
an involved process.
Infozilla~\cite{bettenburg2008extracting}
is based on a set of regular expressions that extract a set of frames
related to a stack trace. The main limitation of this solution is that it is not
able to extract stack traces embedded on verbose log files (i.e., on which we
can find log text mixed with exception frames). Bacchelli
et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
from development emails and relate it to code artifacts (i.e. classes) mentioned
on the stack trace. In addition to those tools, ExceptionMiner is able to 
both extract stack traces from natural language artifacts and to 
classify them in a set of predefined categories.

@@

Additionally, Cabral and Marques~\cite{cabral2007exception} analyzed the
exception handling code of 32 open-source systems, both for Java and .NET. They
observed that the action handlers were very simple (e.g., logging and present a
message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
set of bad practices on exception handling that hinder software maintainability
and robustness, based on their own experience with Java enterprise applications.
Our work differs from those two works as we tried to identify bad practices from
the analysis of stack traces extracted from issues in combination with and bytecode and source
code analysis. \footnote{ to do: TODORESSALTAR LIMITACAO DE QUANTIDADE DE FALSOS POSITIVOS... E QUE OS STACKS SAO REAIS...
REPRESENTAM FLUXOS REAIS.... POST MOSTERN ANALISYS OF REAL EXECUTIONS....}
